---
title: "EDLD 610: Lab 4"
author: "Cameron Kay, Lea Frank, Ashley Miller"
date: "2/5/2019"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

# set knitr options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7.5,
                      fig.height = 4.0)

# install github packages
#devtools::install_github("wilkelab/cowplot")
#install.packages("colorspace", repos = "http://R-Forge.R-project.org")
#devtools::install_github("clauswilke/colorblindr")
#install.packages("rcartocolor")

# load packages
library(tidyverse)
library(magrittr)
library(rio)
library(here)
library(colorblindr)
library(colorspace)
library(cowplot)
library(ggmap)
library(maps)
library(mapdata)
library(ggthemes)
library(rcartocolor)
library(gghighlight)
library(ggrepel)
library(geosphere)
library(paletteer)

# load data
honey_data <- import(here("data/honeyproduction.csv"), setclass = "tibble") %>%
  janitor::clean_names()
```

#### 1) Visualize the total production of honey across years by state. Use color to highlight the west coast (Washington, Oregon, and California). (Ash)

```{r part_1}

honey_data %<>%
  mutate(totalprod_mil = totalprod/1000000)

ggplot(honey_data, aes(x = year, y = totalprod_mil, color = state)) +
  geom_line(size = 1) +
  gghighlight(state == "OR" |
                state == "WA" |
                state == "CA") +
  scale_color_carto_d(palette = "Burg") +
  theme_minimal() +
  theme(axis.title = element_text(colour = "gray20", face = "bold"),
        plot.title = element_text(face = "bold")) +
  labs(title = "Figure 1. Total production of honey (in millions) across years by state",
         subtitle = "Highlighted lines represent west coast states (California, Washington, and Oregon)",
         y = "Total production of honey * 1,000,000",
         x = "Year")

#values = c("#157381", "#4D306E", "#C0A5E1")
#values = c("#fde0dd", "#fa9fb5", "#c51b8a")

```

```{r part_1_alt, include=FALSE, eval=FALSE}

#this approach will also yeild a similar graph to above, but this way is less efficient

to_label <- c("WA", "OR", "CA")
subset_states <- honey_data %>%
    filter(state %in% to_label,
           year == '2012')

ggplot(honey_data, aes(x = year, y = totalprod_mil, group = state)) +
  geom_line(color = "gray20", linetype = 3) +
  geom_line(data = filter(honey_data, 
                          state == "OR" |
                          state == "WA" | 
                          state == "CA"),
            color = "#1CAEAE",
            size = 1) +
  geom_text_repel(aes(label = state), data = subset_states, size = 4) +
  theme_minimal() +
  theme(axis.title = element_text(colour = "gray20"),
        plot.title = element_text(face = "bold")) +
  labs(title = "Figure 1. Total production of honey (in millions) across years by state",
         subtitle = "Highlighted lines represent west coast states (California, Washington, and Oregon)",
         y = "Total production of honey * 1,000,000",
         x = "Year") 

```

#### 2) Reproduce the plot according three different kinds of color blindness, as well as a desaturated version. (Ash)

```{r part_2}

ash_plot1 <- ggplot(honey_data, aes(x = year, y = totalprod_mil, color = state)) +
  geom_line(size = 1) +
  gghighlight(state == "OR" |
                state == "WA" |
                state == "CA") +
  scale_color_carto_d(palette = "Burg") +
  theme_void()

colorblindr::cvd_grid(ash_plot1)

```

Note that the line representing California may not be the best choice if we want the plot to be readable to all viewers (it's hard to see in both the protanomaly and desaturated conditions).

#### 3) Reproduce the plot using a color blind safe pallette. (Ash)

```{r part_3}

ggplot(honey_data, aes(x = year, y = totalprod_mil, color = state)) +
  geom_line(size = 1) +
  gghighlight(state == "OR" |
                state == "WA" |
                state == "CA") +
  scale_color_viridis_d() +
  theme_minimal() +
  theme(axis.title = element_text(colour = "gray20", face = "bold"),
        plot.title = element_text(face = "bold")) +
  labs(title = "Figure 2. Total production of honey (in millions) across years by state",
         subtitle = "Highlighted lines represent west coast states (California, Washington, and Oregon)",
         y = "Total production of honey * 1,000,000",
         x = "Year")
  
```

```{r part_3_color_check}

ash_plot2 <- ggplot(honey_data, aes(x = year, y = totalprod_mil, color = state)) +
  geom_line(size = 1) +
  gghighlight(state == "OR" |
                state == "WA" |
                state == "CA") +
  scale_color_viridis_d() +
  theme_void()

colorblindr::cvd_grid(ash_plot2)

```

While the above plot (Fig. 2) uses a color blind safe pallette, note that the line representing Washington is not easily distinguishable in the desaturated condition.

#### 4. Download the file here denoting the region and division of each state. (Lea)

+ Join the file with your honey file.
+ Produce a bar plot displaying the average honey.
+ Use color to highlight the region of the country the state is from.
+ Note patterns you notice.

```{r part_4_Lea1}
region <- import("https://raw.githubusercontent.com/cphalpert/census-regions/master/us%20census%20bureau%20regions%20and%20divisions.csv")

honey_join <- left_join(honey_data, region, by = c("state" = "State Code"))

honey_join <- honey_join %>% 
  group_by(state) %>% 
  summarize(avgprod = mean(totalprod)) %>% 
  left_join(honey_join)

```

```{r part_4_Lea2, fig.height=6.0}
honey_join <- honey_join %>% 
  ungroup() %>% 
  mutate(avgprod_mil = avgprod/100000)

honey_join %>% 
  ggplot(aes(x = reorder(state, avgprod_mil), y = avgprod_mil, fill = Region)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  scale_fill_OkabeIto() +
  labs(title = "Average Honey Production by State",
       x = "State", y = "Average Honey Production \n (in millions)")
  
```

Something to note from this figure is that the average honey production from 1998 to 2012 was greatest in Midwest and Western states, while the lowest average honey production came out of Northeast and some Southern states. Average honey production in Southern states varied a lot, with Florida producing a large numbers of honey, but Oklahoma and Maryland producing the lowest numbers.

#### 5) Create a heatmap displaying the average honey production across years by region. (Lea)

```{r part_5_LF}
honey_avg_region <- honey_join %>% 
  group_by(Region, year) %>% 
  summarize(avgprod_mil = mean(totalprod)/100000)

cont_colorpal <- paletteer_d("rcartocolor","BurgYl", type = "continuous")

honey_avg_region %>% 
  ggplot(aes(x = year, y = reorder(Region, avgprod_mil), fill = avgprod_mil)) +
  geom_tile() +
  scale_fill_carto_c(palette = "BurgYl",
                     name = "Average Honey Production \n (in millions)") +
  labs(title = "Average Honey Production by Region from 1998 to 2012",
       x = "Year", y = "Region") +
  theme_minimal()
```

#### 6) Create at least one more plot of your choosing using color to distinguish, represent data values, or highlight. If you are interested in producing maps, I would recommend joining the data with the output from ggplot2::map_data("state"). But be careful with keys and that you don’t end up with a many-to-many join. (Cam)

```{r part_6}
# load map data
map_data <- ggplot2::map_data("state")

# join honey_data and map_data
## find unique keys
honey_data %>%
  count(state, year) %>% 
  filter(n > 1)

map_data %>%
  count(order) %>% 
  filter(n > 1)

# summarize data and join
cam_data <- honey_data %>%
  group_by(state) %>%
  summarize(totalprod_m = mean(totalprod, na.rm = TRUE)) %>%
  mutate(totalprod_m = totalprod_m / 1000000) %>%
  rowwise() %>%
  mutate(region = state.name[grep(state, state.abb)],
         region = tolower(region)) %>%
  left_join(map_data, by = "region") %>%
  filter(region != "hawaii")

# check lengths
length(unique(honey_data$state))
length(unique(cam_data$region))

# find centroids
centroids <- data.frame(state = c(unique(cam_data$state)),
                        long  = rep(NA, length(unique(cam_data$state))),
                        lat   = rep(NA, length(unique(cam_data$state))))
                                   
centroid_data <- cam_data %>%
  ungroup() %>%
  select(state, long, lat) 

for (i in 1:length(unique(centroid_data$state))) {
  data_temp <- filter(centroid_data, state == unique(centroid_data$state)[i])
  centroid_temp <- centroid(data_temp[, c("long", "lat")])
  centroids[i, "long"] <- centroid_temp[1]
  centroids[i, "lat"]  <- centroid_temp[2]
}

# plot
ggplot(data = cam_data, aes(x = long, y = lat)) + 
  geom_polygon(data = map_data, aes(group = group), fill = "grey80") +
  geom_polygon(aes(fill = totalprod_m, group = group), color = "white") + 
  coord_fixed(1.3) +
  scale_fill_continuous_sequential("Oranges") +
  labs(title = "Honey production for each of the contiguous states",
       subtitle = "Average production between 1998 and 2012",
       x = "",
       y = "",
       fill = "Total production (millions)") +
  theme_void() +
  theme(legend.position = "bottom") + 
  geom_text(data = centroids, aes(label = state))

```
